Predicting the 2024 U.S. Presidential Election: Trump’s Likely Victory Based on Polling Analysis

### 2024-11-03 paper/paper.qmd
committing changes to data section in paper.qmd and  03-clean_data.R

---
title: "Predicting the 2024 U.S. Presidential Election: Trump’s Likely Victory Based on Polling Analysis"
subtitle: "A Bayesian Approach Reveals Strong Support for Donald Trump Across Key States, Highlighting Critical Trends in Voter Preferences Leading Up to Election Day"
author: 
  - Jimin Lee
  - Sarah Ding
  - Xiyan Chen
thanks: "Code and data are available at: [https://github.com/jamiejiminlee/2024_US_Elections.git](https://github.com/jamiejiminlee/2024_US_Elections.git)."
date: today
date-format: long
abstract: "This paper presents a predictive model for the 2024 United States Presidential Election using state-level polling data, with a Bayesian approach to estimate the outcomes between Donald Trump and Kamala Harris. By focusing on high-quality polls from reputable sources, we aggregated data to predict the likely winner in each state. Our Bayesian hierarchical model, incorporating time-dependent splines and state-specific random effects, captures shifts in candidate support over time. The results underscore Trump’s advantage in critical swing states such as Florida, Ohio, and North Carolina, positioning him as the predicted overall winner of the 2024 election. We discuss the model’s performance, accuracy, and limitations, reflecting on the challenges inherent in electoral forecasting."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction
This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data [@FiveThirtyEight] to forecast the likely winner between Donald Trump and Kamala Harris. We aggregate high-quality poll data to create a Bayesian logistic regression model that estimates the probability of a Trump or Harris victory in each state. By comparing polling percentages for both candidates, we aim to predict election outcomes based on voter support patterns across the states.

The estimand in this study is the probability that Donald Trump or Kamala Harris wins a given state, based on aggregated state-level polling averages. The binary outcome variable in our model indicates whether Trump (1) or Harris (0) is predicted to win in each state, and the predictor variables are the average polling percentages for each candidate.

Our model utilizes a Bayesian approach with a hierarchical structure to predict election outcomes by comparing the average polling percentages for Trump and Harris across states. This methodology incorporates time-dependent splines and state-specific random effects, which allow for capturing variations in support trends over time. Swing states emerge as critical battlegrounds, where polling percentages are closely contested and could influence the final election result.

Accurate election predictions provide valuable insights into voter dynamics and help political analysts, campaigns, and the public anticipate electoral outcomes. By focusing on high-quality polling data, our model enhances the reliability of predictions and highlights key regions where voter sentiment may shift, ultimately affecting the election.

The results of our model predict a victory for Donald Trump in the 2024 U.S. Presidential Election. Key battleground states, such as Florida, Ohio, and North Carolina, lean toward Trump, emphasizing his advantage in critical regions. Our analysis points to Trump as the probable overall winner, with swing states playing a decisive role in determining the final outcome.

The remainder of this paper is structured as follows. In [@sec-data], we describe the data and variables used in the analysis. [@sec-model] outlines the model setup and estimation strategy. [@sec-results] presents the results, including the predicted election outcomes and visualizations. Finally, [@sec-discussion] discusses the implications and limitations of our findings.

# Data {#sec-data}

## Overview

REWRITE

## Measurement

REWRITE

## Outcome variables

REWRITE

```{r}
#| label: fig-averagetable
#| fig-cap: Summary of 2024 U.S. Presidential Election Polling Averages by State for Trump and Harris, with Indicated Leading Candidate. States with missing data, such as Colorado, are labeled as "NA."
#| echo: false
#| warning: false
#| message: false
# Load cleaned data
clean_poll_data <- read_parquet("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.parquet")

@@

# Create winner column: 1 if Trump_pct is greater than Harris_pct, 0 otherwise
state_average <- state_average |>
  mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))

# Prepare table for display with rounding to 1 decimal place
average_table <- state_average |>
  mutate(
    Trump_pct = round(Trump_pct, 1),
    Harris_pct = round(Harris_pct, 1)
  ) |>
  select(
    State = state, 
    "Trump %" = Trump_pct, 
    "Harris %" = Harris_pct, 
    Winner = winner
  ) |>
  mutate(Winner = ifelse(Winner == 1, "Trump", "Harris"))

@@

REWRITE

## Predictor variables

REWRITE

```{r}
#| label: fig-barplot
#| fig-cap: "Comparison of 2024 U.S. Presidential Election Polling Averages for Trump and Harris by State. Bars represent the average polling percentages for each candidate, with states missing data, such as Colorado, labeled accordingly."
#| echo: false
#| warning: false

# Reshape data to a long format for side-by-side bars
stateaverage_long <- state_average %>%
  pivot_longer(cols = c(Trump_pct, Harris_pct),
               names_to = "candidate",
               values_to = "support_pct")

ggplot(stateaverage_long, aes(x = reorder(state, -support_pct), y = support_pct, fill = candidate)) + 
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "black") +
  labs(x = "State", y = "Average Percentage") +
  scale_fill_manual(
    name = "Candidate", 
    values = c("Trump_pct" = "red", "Harris_pct" = "blue"),
    labels = c("Trump_pct" = "Trump %", "Harris_pct" = "Harris %")
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

REWRITE

@@

REWRITE

@@

- \begin{equation}
  \text{Trump\_pct}_i \text{ and } \text{Harris\_pct}_i \text{ are the polling support levels for Donald Trump and Kamala Harris in state } i.
  \end{equation}

- \begin{equation}
  f(\text{end\_date\_num}_i) \text{ is a natural spline function of time (measured as } \text{end\_date\_num}, \text{ with 4 degrees of freedom, capturing the non-linear time trends in support.}
  \end{equation}

- \begin{equation}
  \alpha_{\text{Trump}} \text{ and } \alpha_{\text{Harris}} \text{ are intercepts for the models, representing the baseline support levels for each candidate.}
  \end{equation}

- \begin{equation}
  u_{\text{state}}^{\text{Trump}} \text{ and } u_{\text{state}}^{\text{Harris}} \text{ are state-level random effects, accounting for differences in baseline support across states.}
  \end{equation}

@@

# Generate predictions for each state and date
state_dates <- seq.Date(min(clean_poll_data$end_date), max(clean_poll_data$end_date), by = "1 week")
state_prediction_grid <- expand.grid(state = unique(clean_poll_data$state), end_date = state_dates) %>%
  mutate(end_date_num = as.numeric(end_date - min(clean_poll_data$end_date)))

@@

# Plot state-level support trends over time with adjusted x-axis labels
ggplot(state_predictions, aes(x = end_date)) +
  geom_line(aes(y = Trump_Support, color = "Trump")) +
  geom_line(aes(y = Harris_Support, color = "Harris")) +
  labs(title = "State-Level Support Trends for Trump and Harris Over Time", y = "Predicted Support", x = "Date") +
  facet_wrap(~ state) +
  scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +  # Set date format and break frequency
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    strip.text = element_text(size = 10),  # Control facet label size if needed
    panel.spacing = unit(0.5, "lines"),    # Add spacing between facets
    panel.grid.major = element_line(linewidth = 0.3)  # Adjust gridline thickness
  )

@@

# Predict probability of Trump and Harris winning by comparing posterior predictions for each state
state_predictions <- clean_poll_data %>%
  distinct(state) %>%
  mutate(end_date = election_day, end_date_num = as.numeric(election_day - min(clean_poll_data$end_date)))

@@

# Display state predictions i n table
knitr::kable(state_summary_table)

@@

# Reshape the data for plotting
state_predictions_long <- state_predictions %>%
  select(state, prob_trump_wins, prob_harris_wins) %>%
  pivot_longer(cols = starts_with("prob"), 
               names_to = "Candidate", 
               values_to = "Probability") %>%
  mutate(Candidate = recode(Candidate, 
                            prob_trump_wins = "Trump", 
                            prob_harris_wins = "Harris"))

# Plot double bar graph of predicted winning percentages by state
ggplot(state_predictions_long, aes(x = reorder(state, Probability), y = Probability * 100, fill = Candidate)) +
  geom_col(position = "dodge") +  # Use dodge for side-by-side bars
  coord_flip() +  # Flip coordinates for better readability
  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
  labs(title = "Predicted Winning Percentages by State on Election Day", 
       x = "State", 
       y = "Winning Probability (%)") +
  theme_minimal() +
  theme(legend.position = "right")
```
EDIT - explain graph results

@@

```{r}

#| echo: false
#| eval: true
#| label: fig-us-map
#| fig-cap: "Predicted Winner by State on Election Day."
#| warning: false

# Add predicted winner column based on `prob_trump_wins`
state_predictions <- state_predictions %>%
  mutate(predicted_winner = ifelse(prob_trump_wins > 0.5, "Trump", "Harris"))

# Plot the US map with predicted winner for each state
plot_usmap(data = state_predictions, values = "predicted_winner", regions = "states") +
  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
  labs(title = "Predicted Winner by State on Election Day", fill = "Predicted Winner") +
  theme(legend.position = "right")

>>>>

---
title: "Predicting the 2024 U.S. Presidential Election: Trump’s Likely Victory Based on Polling Analysis"
subtitle: "A Bayesian Approach Reveals Strong Support for Donald Trump Across Key Battleground States, Highlighting Critical Trends in Voter Preferences Leading Up to Election Day"
author: 
  - Jimin Lee
  - Sarah Ding
  - Xiyan Chen
thanks: "Code and data are available at: [https://github.com/jamiejiminlee/2024_US_Elections.git](https://github.com/jamiejiminlee/2024_US_Elections.git)."
date: today
date-format: long
abstract: "This paper presents a predictive model for the 2024 United States Presidential Election, utilizing state-level polling data to focus specifically on key battleground states. By employing a Bayesian approach, we estimate the outcomes between Donald Trump and Kamala Harris, incorporating high-quality polls from reputable sources. Our Bayesian hierarchical model, featuring time-dependent splines and state-specific random effects, effectively captures shifts in candidate support over time. The analysis reveals that Trump is projected to win in critical battleground states such as Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin, highlighting his advantage in these pivotal regions. These findings underscore the significance of battleground states in determining electoral outcomes, particularly in a highly polarized political landscape. We also discuss the model’s performance, accuracy, and limitations, reflecting on the unique challenges inherent in forecasting electoral outcomes in these crucial areas."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction

EDIT & ELABORATE - gpt generated
The 2024 United States Presidential Election presents a unique opportunity to examine the dynamics of voter preferences in key battleground states. With a highly polarized electorate and the stakes at an all-time high, accurately predicting the outcome of this election is crucial for political analysts, campaigns, and the public. This paper develops a predictive model that leverages state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, using a Bayesian hierarchical approach to accommodate variations across states.

The central estimand of this study is the probability of victory for each candidate in battleground states, defined by the binary outcome variable indicating a predicted win for Trump (1) or Harris (0). To achieve this, we utilize aggregated polling percentages as predictor variables, allowing us to capture shifts in voter sentiment over time. By employing logistic regression and accounting for state-specific random effects, our model not only identifies trends in polling data but also highlights critical regions that could determine the overall election outcome.

In our analysis, we focused on key battleground states—Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin—where the margins are often razor-thin and can swing the election in favor of one candidate or another. Our findings indicate that Trump is predicted to win in these crucial states, positioning him as the likely overall victor in the 2024 election. This is significant as it underscores the importance of battleground states in shaping the electoral landscape and highlights the intricate dynamics of voter preferences that can influence the final result.

The structure of this paper is as follows: we begin by detailing the data sources and methodology used for the analysis. Next, we present the results, focusing on the predicted probabilities of each candidate winning in the targeted battleground states. Finally, we discuss the implications of our findings and the limitations of our model, contributing to a better understanding of voter dynamics in the 2024 U.S. Presidential Election.

# Data {#sec-data}
## Overview
This paper utilizes the dataset of the 2024 Presidential Election cycle, obtained from [@FiveThirtyEight]. The dataset is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results. It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc. For every pollster, relevant variables such as state, polling date, transparency score, candidate name, and percentage of support for the appropriate candidate are included. The dataset compiles data from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.

We use the statistical programming language R [cite R]... to perform data simulation, testing, cleaning, and analysis.

## Measurement
In the current presidential election cycle,  Americans’ opinions on voting for their preferred candidate are transformed into data points through a series of steps that involve surveying, data processing, and structuring responses. Pollsters then store these responses in structured databases for analysis. In the dataset obtained from Project 538, each entry in the dataset corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for each candidate. Each pollster has different approaches to sampling respondents, recruiting respondents, as well as questionnaire types, which affects each pollster’s reliability and data quality.

Pollsters begin by selecting a sample from the registered voter population, often using voter files with demographic and contact information. They then apply methods of stratified random sampling to ensure the sample reflects the electorate’s diversity in terms of age, gender, race, education, voter behavior, etc.

Then, pollsters reach out to voters through various channels, such as phone calls, emails, text messages, or online panels. These are called recruiting methods and usually, each of them has its trade-offs. If the respondents are connected through any of these channels, they are asked a series of questions either by a live interviewer or a preset of questions.

The types of questions often include:

- “If the election were held today, would you vote for [Candidate A] or [Candidate B]?” 
- “How strongly do you support [Candidate]?” (e.g., on a scale of 1 to 10, or strongly, somewhat, undecided)
- “Do you support [Policy A] proposed by [Candidate A] or [Policy B] proposed by [Candidate B]?”

Responses are recorded digitally, either by survey software or directly into a database by interviewers, capturing both the answers and relevant demographic data. Each response becomes a data point in the database. For example, if a respondent prefers Candidate A, their answer may be recorded as a binary variable (e.g., “1” for Candidate A, “0” for Candidate B). Along with candidate preference, pollsters capture demographic and behavioral details, such as age, gender, race, education, income, voter history, and party registration.

Each respondent’s answers and demographic details are grouped and saved in a structured database. Pollsters then clean the data by checking for inconsistencies, removing incomplete responses, and addressing non-responses. Pollsters also perform weighting to ensure the representativeness of data collected, details of weighting are further explained in {appendix 1}.

Once cleaned and weighted, data points are aggregated to determine overall candidate support, often producing metrics such as the percentage of respondents supporting each candidate by state or other specific demographics. Each pollster would have a database with all the weighted and aggregated data, and Project 538 presents a compilation of the databases obtained from various pollsters.

## Variables
To give an overview of the data, we provide summary statistics of the key variables, highlighting the characteristics of the polling support for each candidate across the battleground states.
```{r}
#| label: fig-summarystatistics
#| fig-cap: "Vertical Summary Statistics for Trump and Harris Polling Data. This table presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states. All values are rounded to three decimal places, providing a concise overview of the polling landscape as of the selected election date."
#| echo: false
#| warning: false
#| message: false

# Load cleaned data
clean_poll_data <- read_parquet("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.parquet")

# Summary statistics for the relevant columns in clean_poll_data
summary_statistics <- clean_poll_data %>%
  summarize(
    `Total Polls` = n(),
    `Average Trump Support` = round(mean(Trump_pct, na.rm = TRUE), 3),
    `Max Trump Support` = round(max(Trump_pct, na.rm = TRUE), 3),
    `Min Trump Support` = round(min(Trump_pct, na.rm = TRUE), 3),

    `Average Harris Support` = round(mean(Harris_pct, na.rm = TRUE), 3),
    `Max Harris Support` = round(max(Harris_pct, na.rm = TRUE), 3),
    `Min Harris Support` = round(min(Harris_pct, na.rm = TRUE), 3)
  )

# Reshape the summary statistics to a vertical format
summary_statistics_long <- pivot_longer(
  summary_statistics,
  cols = everything(),
  names_to = "Statistic",
  values_to = "Value"
)

# Display vertical summary statistics in a kable table
kable(summary_statistics_long) %>%
  kable_styling(full_width = FALSE, position = "center")

@tbl-summarystatistics reveals a total of 34 polls conducted, with an average support level of 46.321% for Trump and 46.661% for Harris. Trump's support peaked at 51.000%, while the lowest recorded support was 42.500%. For Harris, the maximum support also reached 51.000%, with a minimum of 42.000%. These statistics illustrate the competitive landscape in these critical states, emphasizing the narrow margins that could ultimately influence the election outcome. The data reflects the volatility of voter sentiment leading up to the 2024 Presidential Election, underscoring the importance of these battleground states in determining the final result.

### Outcome Variable
The outcome variable in this study is a binary indicator, specifically constructed to represent the predicted winner of the 2024 U.S. Presidential Election based on polling data from key battleground states. This variable categorizes the election outcome into two distinct categories: a value of "Trump" indicates that Donald Trump is predicted to win, while a value of "Harris" denotes a predicted victory for Kamala Harris. The classification is derived from a conditioning statement that assigns the predicted winner based on the probability of Trump winning; if this probability exceeds 0.5, the predicted winner is labeled as Trump, otherwise, it is labeled as Harris. This binary outcome variable serves a crucial role in facilitating the analysis of voter support dynamics and the effectiveness of the predictive modeling employed in this paper.

By focusing on this dichotomy, we can glean insights into how polling data might translate into electoral outcomes, highlighting the competitive nature of the election in these pivotal states. The construction of the outcome variable not only enhances our understanding of voter preferences but also underscores the significance of binary classification in electoral forecasting, especially in states that could ultimately sway the overall election results.

### Predictive Variable
The predictive variable in this study is the polling support percentage for each candidate, specifically for Donald Trump and Kamala Harris, measured across various battleground states. This variable represents the predicted likelihood of each candidate winning in the specified states based on the polling data collected during the election cycle.
```{r}
#| label: tbl-outcomevar
#| fig-cap: "Average polling support percentages for Donald Trump and Kamala Harris across key battleground states. The table presents the average support levels, indicating Trump’s and Harris’s polling figures in states such as Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada. The comparative analysis highlights the varying levels of candidate support in these critical regions, providing insight into the electoral landscape ahead of the 2024 U.S. Presidential Election."
#| echo: false
#| warning: false
#| message: false

@@

# Prepare table for display with rounding to 1 decimal place
average_table <- state_average |>
  mutate(
    Trump_pct = round(Trump_pct, 1),
    Harris_pct = round(Harris_pct, 1)
  ) |>
  select(
    State = state, 
    "Trump %" = Trump_pct, 
    "Harris %" = Harris_pct, 
  )

@@

```
@tbl-predictivevar highlights the average support levels for both Donald Trump and Kamala Harris across key battleground states, including Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada. The values reflect the polling percentages for each candidate, providing a straightforward comparison of their standings within these critical regions.

Understanding these polling percentages is essential as they offer insight into the electoral landscape and help evaluate which candidate may hold an advantage in the upcoming election. The battleground states depicted in the table are particularly significant due to their historical tendency to swing between parties, making them crucial targets for campaign efforts. Analyzing the outcome variable through these polling percentages informs our predictive modeling and enhances our comprehension of the dynamics at play during the electoral process. This comparison not only contextualizes the current political climate but also underlines the importance of these battleground states in shaping the overall election results. As the campaign progresses, these insights will be invaluable for strategizing and understanding voter sentiment in these pivotal areas.

@@

WRITE SECTION

@@

- \begin{equation}
  \text{Trump\_pct}_i \text{ and } \text{Harris\_pct}_i \text{ are the polling support levels for Donald Trump and Kamala Harris in state } i.
  \end{equation}
- \begin{equation}
  f(\text{end\_date\_num}_i) \text{ is a natural spline function of time (measured as } \text{end\_date\_num}, \text{ with 4 degrees of freedom, capturing the non-linear time trends in support.}
  \end{equation}
- \begin{equation}
  \alpha_{\text{Trump}} \text{ and } \alpha_{\text{Harris}} \text{ are intercepts for the models, representing the baseline support levels for each candidate.}
  \end{equation}
- \begin{equation}
  u_{\text{state}}^{\text{Trump}} \text{ and } u_{\text{state}}^{\text{Harris}} \text{ are state-level random effects, accounting for differences in baseline support across states.}
  \end{equation}

@@

# Define battleground states
battleground_states <- c("Arizona", "Georgia", "Michigan", "Nevada", 
                          "North Carolina", "Pennsylvania", "Wisconsin")

# Generate predictions for each state and date
state_dates <- seq.Date(min(clean_poll_data$end_date), max(clean_poll_data$end_date), by = "1 week")
state_prediction_grid <- expand.grid(state = battleground_states, end_date = state_dates) %>%
  mutate(end_date_num = as.numeric(end_date - min(clean_poll_data$end_date)))

@@

# Plot state-level support trends over time with adjusted x-axis labels
ggplot(state_predictions, aes(x = end_date)) +
  geom_line(aes(y = Trump_Support, color = "Trump")) +
  geom_line(aes(y = Harris_Support, color = "Harris")) +
  labs(title = "State-Level Support Trends for Trump and Harris Over Time (Battleground States)", 
       y = "Predicted Support", 
       x = "Date") +
  facet_wrap(~ state) +
  scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +  # Set date format and break frequency
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    strip.text = element_text(size = 10),  # Control facet label size if needed
    panel.spacing = unit(0.5, "lines"),    # Add spacing between facets
    panel.grid.major = element_line(linewidth = 0.3)  # Adjust gridline thickness
  )

@@

# Predict probability of Trump and Harris winning by comparing posterior predictions for each state
state_predictions <- clean_poll_data %>%
  distinct(state) %>%
  filter(state %in% battleground_states) %>%  # Only include battleground states
  mutate(end_date = election_day, 
         end_date_num = as.numeric(end_date - min(clean_poll_data$end_date)))

@@

# Display state predictions in table
knitr::kable(state_summary_table) %>%
  kable_styling(full_width = FALSE, position = "center")

@@

# Reshape the data for plotting only for battleground states
state_predictions_long <- state_predictions %>%
  select(state, prob_trump_wins, prob_harris_wins) %>%
  filter(state %in% battleground_states) %>%  # Ensure we're only working with battleground states
  pivot_longer(cols = starts_with("prob"), 
               names_to = "Candidate", 
               values_to = "Probability") %>%
  mutate(Candidate = recode(Candidate, 
                            prob_trump_wins = "Trump", 
                            prob_harris_wins = "Harris"))

# Plot double bar graph of predicted winning percentages by state
ggplot(state_predictions_long, aes(x = reorder(state, Probability), y = Probability * 100, fill = Candidate)) +
  geom_col(position = "dodge") +  # Use dodge for side-by-side bars
  coord_flip() +  # Flip coordinates for better readability
  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
  labs(title = "Predicted Winning Percentages by Battleground State on Election Day", 
       x = "State", 
       y = "Winning Probability (%)") +
  theme_minimal() +
  theme(legend.position = "right")

```
EDIT - explain graph results

@@

```{r}
#| echo: false
#| eval: true
#| label: fig-us-map
#| fig-cap: "Predicted Winner by Battleground State on Election Day."
#| warning: false

# Add predicted winner column based on `prob_trump_wins`
state_predictions <- state_predictions %>%
  filter(state %in% battleground_states) %>%  # Ensure we only include battleground states
  mutate(predicted_winner = ifelse(prob_trump_wins > 0.5, "Trump", "Harris"))

# Plot the US map with predicted winner for each state
plot_usmap(data = state_predictions, values = "predicted_winner", regions = "states") +
  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
  labs(title = "Predicted Winner by Battleground State on Election Day", fill = "Predicted Winner") +
  theme(legend.position = "right")
